---
title: "R Notebook"
output: html_notebook
---

```{r}

rm(list = ls())

suppressMessages(library(data.table))
suppressMessages(library(tidyverse))
suppressMessages(library(lubridate))
suppressMessages(library(stringr))

```

## Get data into R

```{r}

orders_orig <- data.frame(fread("data/orders.csv", header = T))
order_items <- data.frame(fread("data/order-items.csv", header = T))

```

## Create separate data frames for orders, customers, items

```{r}

customers <- orders_orig %>%
  select(customers_id, customers_gender, customers_dob) %>%
  unique() %>%
  mutate(customers_id = as.numeric(customers_id)) %>%
  arrange(customers_id)

orders <- orders_orig %>%
  select(-c(customers_gender, customers_dob)) %>%
  unique() %>%
  mutate(customers_id = as.numeric(customers_id)) %>%
  arrange(customers_id)


```

## Clean the order items table

```{r}

order_items <- order_items %>%
  mutate(products_price = as.numeric(products_price)) %>%
  mutate(products_quantity = as.numeric(products_quantity)) %>%
  filter(products_price > 0) %>%
  mutate(products_quantity = ifelse(str_detect(products_name, "case|CASE|Case") & !str_detect(products_name, "6 bottles"), products_quantity * 12, products_quantity)) %>%
  mutate(products_quantity = ifelse(str_detect(products_name, "case|CASE|Case") & str_detect(products_name, "6 bottles"), products_quantity * 6, products_quantity)) %>%
  mutate(products_quantity = ifelse(str_detect(products_name, "12 Assorted"), products_quantity * 12, products_quantity)) %>%
  mutate(products_quantity = ifelse(str_detect(products_name, "Assorted"), products_quantity * 12, products_quantity)) %>%
  mutate(products_quantity = ifelse(str_detect(products_name, "Budget White"), products_quantity * 12, products_quantity)) %>%
  mutate(products_quantity = ifelse(str_detect(products_name, "30 bottles"), products_quantity * 30, products_quantity)) %>%
  mutate(products_quantity = ifelse(str_detect(products_name, "Voucher"), products_quantity * 6, products_quantity))


# TODO: 

# mixed cases contain 12 bottles (6 red, 6 white)
# seems all order items with "case" in the name, and not a multiple of 6 (or 12?) are actually cases and should be multiplied by 12
# assuming assorted is 12 too
# wynklub seems to be just one bottle
# gift vouchers are tricky ... assume the intention is to buy 6 bottles
# remove "as arranged by Johan"?
# remove wine years?

# table(order_items$products_quantity)
# order_items[order_items$products_quantity == 1 & !str_detect(order_items$products_name, "wynklub"),]
# order_items[str_detect(order_items$products_name, "Johan"),]

```

## Add features to customers table

```{r}
# number of orders

customers <- orders %>%
  count(customers_id) %>%
  right_join(customers)

colnames(customers)[colnames(customers) == "n"] <- "order_count"

# age and age group

customers <- customers %>%
  mutate(customers_dob = parse_date_time(customers_dob, "%Y-%m-%d H:%M:%S")) %>%
  mutate(customer_age =  as.period(interval(customers_dob, ymd("20170824")))) %>%
  mutate(customer_age_group = ifelse(customer_age >= years(60), "old", ifelse(customer_age >= years(35), "mid", "young"))) %>%
  mutate(customer_age_group = as.factor(customer_age_group))

# avg. per bottle
customer_item_prices <- left_join(orders, order_items) %>%
  mutate(total_paid = products_price * products_quantity) %>%
  group_by(customers_id) %>%
  summarise("order_amount" = sum(total_paid), "item_count" = sum(products_quantity)) %>%
  mutate(avg_item = order_amount / item_count)

customers <- customer_item_prices %>%
  select(customers_id, avg_item) %>%
  right_join(customers)

# customers %>%
#   arrange(desc(order_count))

```

Identify top and bottom customers by number of unique products bought

```{r}

# justify choosing "39263" for a top customers
order_items %>%
  right_join(orders, by = "orders_id") %>%
  count(customers_id) %>%
  arrange(desc(n)) %>%
  head()

# justify choosings "411" for a bottom customer
order_items %>%
  right_join(orders, by = "orders_id") %>%
  count(customers_id) %>%
  arrange(n) %>%
  head()


```


## Simple bought products approach (i.e. user-based collaborative filtering)

```{r}

# get matrix of "bought wines" i.e. customer_id as rows and wines as columns

customer_products_bought_tall <- order_items %>%
  select(orders_id, products_name) %>%
  inner_join(orders %>% select(customers_id, orders_id)) %>% # use inner join to avoid customers without purchases?
  select(customers_id, products_name) %>%
  unique() %>%
  mutate(bought = 1)

customer_products_bought_wide <- customer_products_bought_tall %>%
  complete(customers_id, products_name, fill = list(bought = 0)) %>%
  spread(key = products_name, value = bought)

# convert data to matrix form 
sorted_customers_id <- as.character(unlist(customer_products_bought_wide[,1]))
customer_products_bought_wide <- as.matrix(customer_products_bought_wide[,-1])
row.names(customer_products_bought_wide) <- sorted_customers_id
```

```{r, eval=F}

cosine_sim <- function(a,b){crossprod(a,b)/sqrt(crossprod(a)*crossprod(b))}
temp_func <- Vectorize(function(x, y) cosine_sim(customer_products_bought_wide[x,], customer_products_bought_wide[y,]))

customer_similarities1 <- outer(seq_len(nrow(customer_products_bought_wide)),
      seq_len(nrow(customer_products_bought_wide)),
      temp_func)

diag(customer_similarities1) <- 0
row.names(customer_similarities1) <- row.names(customer_products_bought_wide)
colnames(customer_similarities1) <- row.names(customer_products_bought_wide)

saveRDS(customer_similarities1, "output/customer_similarities1.rds")

# max(customer_products_bought_wide[263,])
# cosine_sim(customer_products_bought_wide[3,], customer_products_bought_wide[3,])

```
```{r}
customer_similarities1 <- readRDS("output/customer_similarities1.rds")


# sort(apply(customer_products_bought_wide, 1, sum), decreasing = T)
# # Raka Spliced 2014
# # Le Bonheur Prima 2013 unlabelled
# # Knorhoek Shiraz/Cabernet Franc/Cabernet Sauvignon 2015
# 
# crossprod(customer_products_bought_wide[, "Raka Spliced 2014"], customer_similarities1["39263",])
# 
# temp <- customer_similarities1["39263",] %*% customer_products_bought_wide
# # customer_similarities1["39263",] %*% customer_products_bought_wide
# dim(customer_similarities1["39263",])
dim(customer_products_bought_wide)
# dim(temp)

dim(customer_products_bought_wide[0,])

customer_scores <- data.frame(product = colnames(customer_products_bought_wide),
                              score = as.vector(customer_similarities1["39263",] %*% customer_products_bought_wide),
                              bought = customer_products_bought_wide["39263",])

customer_scores %>%
  # filter(bought == 0) %>%
  arrange(desc(score)) %>%
  select(product, score, bought)
```

```{r}

customer_based_recommendations <- function(customer, customer_similarities, bought_wines){
  
  # turn into character if not already
  customer <- ifelse(is.character(customer),customer,as.character(customer))
  
  # get scores
  customer_scores <- data.frame(product = colnames(bought_wines), 
                            score = as.vector(customer_similarities[customer,] %*% bought_wines), 
                            bought = bought_wines[customer,])
  
  # sort unseen movies by score and remove the 'seen' column
  customer_recom <- customer_scores %>% 
    filter(bought == 0) %>%
    arrange(desc(score)) %>% 
    select(-bought) 
  
  return(customer_recom)
  
}

customer_based_recommendations("39263", customer_similarities1, customer_products_bought_wide)

```

## Item-based approach

```{r, eval=F}

products_customers_wide <- t(customer_products_bought_wide)

temp_func <- Vectorize(function(x, y) cosine_sim(products_customers_wide[x,], products_customers_wide[y,]))

product_similarities1 <- outer(seq_len(nrow(products_customers_wide)),
      seq_len(nrow(products_customers_wide)),
      temp_func)

diag(product_similarities1) <- 0
row.names(product_similarities1) <- row.names(products_customers_wide)
colnames(product_similarities1) <- row.names(products_customers_wide)

saveRDS(product_similarities1, "output/product_similarities1.rds")
```
```{r}
product_similarities1 <- readRDS("output/product_similarities1.rds")
```
```{r}
user_bought <- customer_products_bought_tall %>% 
        filter(customers_id == "39263") %>% 
        select(products_name) %>% 
        unlist() %>% 
        as.character()

product_scores <- tibble(product = row.names(product_similarities1), 
                      score = apply(product_similarities1[,user_bought],1,sum),
                      bought = products_customers_wide[,as.character(39263)])

product_scores %>%
  filter(bought == 0) %>%
  arrange(desc(score))

```
```{r}
product_based_recommendations <- function(customer, product_similarities, bought_products){
  
  # turn into character if not already
  customer <- ifelse(is.character(customer),customer,as.character(customer))
  
  # get scores
  customer_bought <- row.names(product_similarities)[bought_products[,customer] == TRUE]
  customer_scores <- tibble(product = row.names(product_similarities), 
                        score = apply(data.frame(product_similarities[,customer_bought]),1,sum),
                        bought = bought_products[,customer])
  # sort unseen movies by score and remove the 'seen' column
  customer_recom <- customer_scores %>% filter(bought == 0) %>% arrange(desc(score)) %>% select(-bought)

  return(customer_recom)
  
}

product_based_recommendations("411", product_similarities1, products_customers_wide)

# undebug(product_based_recommendations)
# test <- "411"
# test <- row.names(product_similarities1)[products_customers_wide[,test] == TRUE]
# tibble(product = row.names(product_similarities1), 
#                         score = apply(data.frame(product_similarities1[,test]),1,sum),
#                         bought = products_customers_wide[,"411"]) %>%
#   arrange(desc(score))


# head(product_based_recommendations("39263", product_similarities1, products_customers_wide)) %in% head(customer_based_recommendations("39263", customer_similarities1, customer_products_bought_wide))


```

## Matrix factorization

Proxy for wine rating? Number of bottles?

Some combination of quantity and price ... i.e. something may be really good, but expensive. Something may be mediocre but cheap. Number of bottles bias towards cheap, price per bottle bias toward expensive. Some combination of the two? And probably a log of that combination in order to reduce impact of MASSIVE orders.

```{r}
product_ratings <- order_items %>%
  mutate(log_order_val = log(products_quantity * products_price)) %>% # create ratings proxy
  inner_join(orders) %>%
  select(customers_id, products_name, log_order_val) %>%
  group_by(customers_id, products_name) %>%
  summarise(rating_proxy = mean(log_order_val)) %>%
  ungroup() %>%
  # complete(customers_id, products_name, fill = list(ratings_proxy = NA))
  complete(customers_id, products_name) %>%
  spread(key = products_name, value = rating_proxy)

# product_ratings %>%
#   gather(-customers_id, key = products_name, value = rating_proxy)

products_list <- as.character(unlist(product_ratings[,1]))
product_ratings <- as.matrix(product_ratings[,-1])
row.names(product_ratings) <- products_list

```


```{r, eval=F}

ranks <- c(1, 2, 3, 4, 5, 10, 20, 50, 100, 200)
# ranks <- c(1, 2, 3)
mf_acc <- NULL
mf_test_acc <- NULL

set.seed(1)
pr_sample_ind <- sample(length(product_ratings), ceiling(0.2*length(product_ratings)))
pr_train <- cbind(product_ratings)
pr_train[pr_sample_ind] <- NA
pr_test <- cbind(product_ratings)
pr_test[-pr_sample_ind] <- NA

for (rank in ranks){
  set.seed(1)

  decomp <- nnmf(A = pr_train,
                 alpha = 3e-3,
                 beta = 3e-3,
                 method = "scd",
                 k = rank,
                 max.iter = 10000
                 )
  mf_observed <- decomp$W %*% decomp$H
  
  # Train accuracy
  mf_train_errors <- (mf_observed - pr_train)^2
  mf_train_accuracy <- sqrt(mean(mf_train_errors[!is.na(pr_train)]))
  
  mf_acc <- append(mf_acc, mf_train_accuracy)
  
  # Test accuracy
  mf_test_errors <- (mf_observed - pr_test)^2
  mf_test_accuracy <- sqrt(mean(mf_test_errors[!is.na(pr_test)]))
  
  mf_test_acc <- append(mf_test_acc, mf_test_accuracy)
  
}

rank_test_results <- data.frame("Rank" = ranks, "Train_acc" = mf_acc, "Test_acc" = mf_test_acc)

saveRDS(rank_test_results, "output/rank_test_results.rds")
```
```{r}


check_matrix <- matrix(rep(mean(pr_train, na.rm = T), length(pr_test)), dim(pr_test)[1], dim(pr_test)[2])
check_errors <- (check_matrix - pr_test)^2
check_accuracy <- sqrt(mean(check_errors[!is.na(pr_test)]))
check_accuracy


```
```{r}
# rank_test_results <- readRDS("output/rank_test_results.rds")

ggplot(rank_test_results) +
  geom_line(aes(x = Rank, y = Train_acc, color = "Train")) +
  geom_line(aes(x = Rank, y = Test_acc, color = "Test")) +
  geom_vline(aes(xintercept = 1), linetype = 2) +
  # scale_color_manual(values = c("Train", "Test"))
  scale_color_discrete("Accuracy") +
  theme_light()

sum(is.na(product_ratings)) / length(product_ratings)

```
```{r, eval=F}
library(NNLM)

set.seed(7)

mf1_decomp <- nnmf(A = product_ratings,
                  alpha = 3e-3,
                  beta = 3e-3,
                  method = "scd",
                  k = 1,
                  max.iter = 10000
                  )

saveRDS(mf1_decomp, "models/mf1_decomp.rds")
```
```{r}

# Top choices for our user
mf1_predicted <- mf1_decomp$W %*% mf1_decomp$H
mf1_predicted[0,head(order(mf1_predicted["411",], decreasing = T))]

# Top wines based on H factor
mf1_predicted[0,head(order(mf1_decomp$H, decreasing = T))]

```



## Customer similarities with user data
```{r}

# get matrix of "bought wines" i.e. customer_id as rows and wines as columns

customer_products_bought_tall_2 <- order_items %>%
  select(orders_id, products_name) %>%
  inner_join(orders %>% select(customers_id, orders_id)) %>% # use inner join to avoid customers without purchases?
  select(customers_id, products_name) %>%
  unique() %>%
  mutate(bought = 1)

# create customer attributes data to add
# customers %>%
#   select(customers_id, customers_gender, customer_age_group) %>%
#   mutate(bought = 1) %>%
#   complete(customers_id, customers_gender, customer_age_group, fill = list(bought = 0))

customer_products_bought_tall_2 <- customers %>%
  select(customers_id, customer_age_group) %>%
  mutate(bought = 1) %>%
  complete(customers_id, customer_age_group, fill = list(bought = 0)) %>%
  rename(products_name = customer_age_group) %>%
  rbind(customer_products_bought_tall_2)

customer_products_bought_tall_2 <- customers %>%
  select(customers_id, customers_gender) %>%
  mutate(bought = 1) %>%
  complete(customers_id, customers_gender, fill = list(bought = 0)) %>%
  rename(products_name = customers_gender) %>%
  rbind(customer_products_bought_tall_2)

customer_products_bought_wide_2 <- customer_products_bought_tall_2 %>%
  complete(customers_id, products_name, fill = list(bought = 0)) %>%
  spread(key = products_name, value = bought)

# convert data to matrix form 
sorted_customers_id <- as.character(unlist(customer_products_bought_wide_2[,1]))
customer_products_bought_wide_2 <- as.matrix(customer_products_bought_wide_2[,-1])
row.names(customer_products_bought_wide_2) <- sorted_customers_id
```

```{r, eval=F}

# cosine_sim <- function(a,b){crossprod(a,b)/sqrt(crossprod(a)*crossprod(b))}


cosine_sim <- function(a,b){crossprod(a,b)/sqrt(crossprod(a)*crossprod(b))}
temp_func_2 <- Vectorize(function(x, y) cosine_sim(customer_products_bought_wide_2[x,], customer_products_bought_wide_2[y,]))

customer_similarities2 <- outer(seq_len(nrow(customer_products_bought_wide_2)),
      seq_len(nrow(customer_products_bought_wide_2)),
      temp_func_2)

diag(customer_similarities2) <- 0
row.names(customer_similarities2) <- row.names(customer_products_bought_wide_2)
colnames(customer_similarities2) <- row.names(customer_products_bought_wide_2)

saveRDS(customer_similarities2, "output/customer_similarities2.rds")

# max(customer_products_bought_wide[263,])
# cosine_sim(customer_products_bought_wide[3,], customer_products_bought_wide[3,])

```
```{r}

head(customer_based_recommendations("39263",
                                    customer_similarities1,
                                    customer_products_bought_wide))

cs_2_indexedBy1 <- customer_similarities2[row.names(customer_similarities1),colnames(customer_similarities1)]
head(customer_based_recommendations("39263",
                                    cs_2_indexedBy1,
                                    customer_products_bought_wide))

```
  
## Combine recommendations

```{r}

ensemble_recommender <- function(customer, context_similarities, no_context_similarities, product_similarities, wide_customer_products_matrix, predicted_ratings){
  
  # Convert to character
  customer <- ifelse(is.character(customer), customer, as.character(customer))
  
  # Context containing similarity
  rank_1 <- customer_based_recommendations(customer,
                                      context_similarities,
                                      wide_customer_products_matrix) %>%
    mutate(p_rank = rank(-score)) %>%
    select(product, p_rank)
  
  # User bought similarity only
  rank_2 <- customer_based_recommendations(customer,
                                      no_context_similarities,
                                      wide_customer_products_matrix) %>%
    mutate(p_rank = rank(-score)) %>%
    select(product, p_rank)
  
  # Item similarity
  rank_3 <- product_based_recommendations(customer,
                                     product_similarities,
                                     t(wide_customer_products_matrix)) %>%
    mutate(p_rank = rank(-score)) %>%
    select(product, p_rank)
  
  # Matrix factorization
  rank_4 <- data.frame(score = predicted_ratings[customer,]) %>%
    rownames_to_column("product") %>%
    mutate(p_rank = rank(-score)) %>%
    select(product, p_rank)
  
  # Combination of methods
  customer_recs <- inner_join(inner_join(inner_join(rank_1, rank_2, by = "product"), rank_3, by = "product"), rank_4, by = "product") %>%
    mutate(overall = p_rank.x + p_rank.y + p_rank.x.x + p_rank.y.y) %>%
    select(product, overall) %>%
    mutate(p_rank = rank(overall)) %>%
    select(product, p_rank) %>%
    arrange(p_rank)
  
  return(customer_recs)
}

suppressWarnings(ensemble_recommender("624",
                     context_similarities = cs_2_indexedBy1,
                     no_context_similarities = customer_similarities1,
                     product_similarities = product_similarities1,
                     wide_customer_products_matrix = customer_products_bought_wide,
                     predicted_ratings = mf1_predicted))

# debug(ensemble_recommender)
# undebug(ensemble_recommender)

# test <- "411"
# data.frame(score = mf1_predicted[test,]) %>%
#     rownames_to_column("product") %>%
#     mutate(p_rank = rank(-score)) %>%
#     select(product, p_rank)

# test <- "411"
# product_based_recommendations(test,
#                                      product_similarities1,
#                                      t(customer_products_bought_wide)) %>%
#     mutate(p_rank = rank(-score)) %>%
#     select(product, p_rank)

# customer_similarities1["411",] %>% data.frame() %>% rownames_to_column() %>% filter(. > 0) # customer exists and has non-zero similarities to other customers
# customer_products_bought_wide["411",] %>% data.frame() %>% rownames_to_column() %>% arrange(desc(.)) # only one product with a non-zero
# cs_2_indexedBy1["411",] %>% data.frame() %>% rownames_to_column() %>% filter(. > 0) # customer exists and has non-zero similarities to other customers
# mf1_predicted["411",] %>% data.frame() %>% rownames_to_column() %>% filter(. > 0) # customer exists and has non-zero ratings

# customer_products_bought_wide["411",] %>% data.frame() %>% rownames_to_column() %>% arrange(desc(.)) # only one product with a non-zero
# product_similarities1["TOP SELLING July Everyday Case",]  %>% data.frame() %>% rownames_to_column() %>% arrange(desc(.)) # product bought exists and has non-zero similarities to other products

```